---
title: "TidyTuesday - Animal Crossing"
author: "Usman Khaliq"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
# Libraries
library(tidyverse)
library(readr)

# Parameters

#url for critic data
url_critic <- "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-05/critic.tsv"

#url for user reviews
url_user_reviews <- "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-05/user_reviews.tsv"

#url for items 
url_items <- "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-05/items.csv"

#url for villagers
url_villagers <- "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-05/villagers.csv"

#===============================================================================

# Code 

#load the data

critic <- 
  url_critic %>% 
  read_tsv()

user_reviews <-
  url_user_reviews %>% 
  read_tsv()

items <-
  url_items %>% 
  read_tsv()

villagers <-
  url_villagers %>% 
  read_tsv()

```

## Start!

I haven't bought Animal Crossing on my Switch yet; I have discovered a newfound addiction for the Legend of Zelda: Breadth of the Wild, and have vowed to finally complete it this year. However, I am happy to dig into this TidyTuesday dataset, and explore this dataset from two angles right now:

* Trends in User Reviews and Critic Reviews
* Sentiment analysis on user's opinions about the game, and exploring whether there are any general differences in how users and critics are reviewing the game.

## Explore the data

First, lets explore the animal crossing dataset

```{r}
glimpse(critic)
summary(critic)
```

```{r} 
glimpse(user_reviews)
summary(user_reviews)
```

## Is Animal Crossing the New Darling of the Quarantined Critics? 

Lets see if critics unanimously love Animal Crossing

```{r} 

critic %>%
  count(grade) %>% 
  arrange(desc(n)) %>% 
  mutate(grade = fct_reorder(factor(grade), n, .desc = TRUE)) %>% 
  ggplot(aes(grade, n)) +
  geom_col() +
  labs(
    title = "How Did Critics Rate Animal Crossing?",
    subtitle = "Animal Crossing is one of the highest rated Games of the Pandemic Year so far",
    x = "Critic Rating (Out of 100)",
    y = "Number of Critics",
    caption = "Source: VillagerDB and MetaCritic, courtesy TidyTuesday"
  )
```

## Do Fans Also Love Animal Crossing With the Same Dedication?

Lets see if fans also are as gagga over the game as the critics are.

```{r}
user_reviews %>%
  count(grade) %>% 
  arrange(desc(n)) %>% 
  mutate(grade = fct_reorder(factor(grade), n, .desc = TRUE)) %>% 
  ggplot(aes(grade, n)) +
  geom_col() +
  labs(
    title = "How Did Users? Rate Animal Crossing?",
    subtitle = "Surprisingly the majority of Users Gave it a 0!",
    x = "User Rating (Out of 100)",
    y = "Number of Users",
    caption = "Source: VillagerDB and MetaCritic, courtesy TidyTuesday"
  )

```


