---
title: "Volcano Eruptions"
author: "Usman Khaliq"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
# Libraries
library(tidyverse)
library(readr)
library(tidymodels)
library(themis)
library(workflows)
library(tune)
library(vip)
library(janitor)

# Parameters
# Parameters

#volcano link
volcano_url <- "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-12/volcano.csv"

#eruptions link
eruptions_url <- "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-12/eruptions.csv"

#events link
events_url <- "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-12/events.csv"

#tree rings link
tree_rings_url <- "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-12/tree_rings.csv"

#sulfur link
sulfur_url <- "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-12/sulfur.csv" 

volcano <-
  volcano_url %>% 
  read_csv()

eruptions <- 
  eruptions_url %>% 
  read_csv()

events <- 
  events_url %>% 
  read_csv()

tree_rings <- 
  tree_rings_url %>% 
  read_csv()

sulfur <-
  sulfur_url %>% 
  read_csv()

#===============================================================================

# Code
```

Lets build a multiclass random forest classifier to predict the type of volcano based on other volcanic characteristics like latitude, tectonic settings etc. 

## Reference 

I referred to Julia Silge's excellent [tutorial](https://www.youtube.com/watch?v=0WCmLYvfHMw&list=LL-xYEpkLu4daSJnUSohR-pw&index=2&t=0s) on building multiclass classifiers with tidymodels.

### Explore the data

```{r}
volcano %>% 
  count(primary_volcano_type, sort = TRUE)
```

Since there are 26 different categories of volcanoes in the dataset and less than a 1000 rows in the `volcano` dataset, lets try to predict a lower number of volcano types.

```{r}
volcano_df <-
  volcano %>% 
  transmute(
    volcano_type = 
      case_when(
        str_detect(primary_volcano_type, "Stratovolcano") ~ "Stratovolcano",
        str_detect(primary_volcano_type, "Shield")        ~ "Shield",
        TRUE                                              ~ "Other"
      ),
    volcano_number,
    latitude,
    longitude,
    elevation,
    tectonic_settings,
    major_rock_1
  ) %>% 
  mutate_if(is_character, factor)
```

## Exploratory Data Analysis 

```{r}
world <- map_data("world") 

ggplot() +
  geom_map(
    data = world, 
    map = world,
    aes(
      long,
      lat,
      map_id = region
    ),
    colour = "white",
    fill = "gray50",
    alpha = 0.2
  ) +
  geom_point(
    data = volcano_df,
    aes(longitude, latitude, color = volcano_type),
    alpha = 0.8
  )
```

The above visualization gives an interesting distribution about the different types of volcanoes around the world. 

## Build a Model 

```{r}
volcano_boot <- 
  volcano_df %>% 
  bootstraps()

```

Lets observe how many volcano categories there are in our data frame.

```{r}
volcano_df %>% 
  count(volcano_type, sort = TRUE)
```


As we can see from the above, the number of Shield volcano types are lower than the other two categories. Lets do some data cleaning to fix this by using the `themis` package that explicitly deals with class imbalance. We will be using the `step_smote` function, that uses the SMOTE algorithm to generate new examples of the minority class using nearest neighbours of the cases. To do this, we would need to create dummy variables for our factor variables, and also centre the numeric variables in `volcano_df`. 


### Preprocessing

```{r}
volcano_df %>% 
  count(tectonic_settings, sort = TRUE)
```

We can also see from above that there are around 11 `tectonic_settings`, so lets collapse a few of the smaller categories of tectonic_settings. 

```{r}
volcano_df %>% 
  count(major_rock_1, sort = TRUE)
```

We can also collapse a few of the categories of the rocks. 


```{r}
volcano_rec <-
  recipe(volcano_type ~ ., data = volcano_df) %>% 
  update_role(volcano_number, new_role = "Id") %>% 
  step_other(tectonic_settings) %>% 
  step_other(major_rock_1) %>% 
  step_dummy(tectonic_settings, major_rock_1) %>% 
  step_zv(all_predictors()) %>% 
  step_normalize(all_predictors()) %>% 
  step_smote(volcano_type)

volcano_prep <-
  volcano_rec %>% 
  prep()
  
```

To look at the results from a recipe, we can `juice` from it.

```{r}
juice(volcano_prep)
```

### Setting up the Model.

First, lets create the specifications for a random forest model. 

```{r}
rf_spec <-
  rand_forest(trees = 1000) %>% 
  set_mode("classification") %>% 
  set_engine("ranger")
```

Lets create a modelling workflow that can hold together the different parts of our random forest model. 

```{r}
volcano_wf <-
  workflow() %>% 
  add_recipe(volcano_rec) %>% 
  add_model(rf_spec)
  
```

By setting `save_pred = TRUE` in the `fit_resamples` function below, we can save and check which one of the volcanoes were correctly predicted by our model. 

```{r}
volcano_res <-
  fit_resamples(
  volcano_wf,
  resamples = volcano_boot,
  control = control_resamples(save_pred = TRUE)
)
```

### Explore the Results 

```{r}
volcano_res %>% 
  collect_metrics()
```

```{r}
volcano_res %>% 
  collect_predictions() %>% 
  conf_mat(volcano_type, .pred_class)
```

```{r}
volcano_res %>% 
  collect_predictions() %>% 
  ppv(volcano_type, .pred_class)
```

```{r}
volcano_res %>% 
  collect_predictions() %>% 
  group_by(id) %>% 
  ppv(volcano_type, .pred_class) %>% 
  ggplot(aes(.estimate)) +
  geom_histogram()
```

Now, lets understand how different variables have performed in our random forest model by using the `vip` package.

```{r}
rf_spec %>% 
  set_engine("ranger", importance = "permutation") %>% 
  fit(
    volcano_type ~ .,
    data = 
      juice(volcano_prep) %>% 
      select(-volcano_number) %>% 
      janitor::clean_names()
  ) %>% 
  vip(geom = "point")
  
```

The biggest variables impacting prediction are latitude and longitude. 

Now, lets make another map with the bootstrapped samples. 

```{r}
volcano_pred <- 
  volcano_res %>% 
  collect_predictions() %>% 
  mutate(correct = volcano_type == .pred_class) %>% 
  left_join(
    volcano_df %>% 
      mutate(.row = row_number())
  )
```

In `volcano_pred`, we have the location of every bootstrapped volcano and whether the prediction was correct or not for its type. 


Finally, lets make a map of `volcano_pred`

```{r}
ggplot() +
  geom_map(
    data = world, 
    map = world,
    aes(
      long,
      lat,
      map_id = region
    ),
    colour = "white",
    fill = "gray50",
    alpha = 0.2
  ) +
  stat_summary_hex(
    data = volcano_pred,
    aes(longitude, latitude, z = as.integer(correct)),
    fun = "mean",
    alpha = 0.7,
    bins = 70
  ) +
  scale_fill_gradient(
    high = "cyan3",
    labels = scales::percent
  ) +
  labs(
    title = "Distribution of Volcanoes and their Prediction Accuracy",
    fill = "Percent classified \ncorrectly"
  )
  

```

In the above visualization, we can see where in the world our random forest model was more correct in predicting the volcano type(the ligher blue regions) and where it was less correct in predicting the volcano type(the dark blue regions).

We can mak the above map better by using a better projection. 

Note:
* in multiclass classification, our metrics for measuring prediction need to be carefully decided. 

